---
title: "Group_9_Analysis"
author: "Group 9"
date: "2023-03-09"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Our group's data is IMDB film database. This database contains a variety of information on all films that have been released. And our research target is to analyse the properties that influence rate of movies. 

```{r loadpackages,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# Load required packages
library(tidyverse)
library(ggplot2)
library(scales)
library(dplyr)
library(broom)
library(pROC)
```

# Step 1: Data pre-processing

```{r data, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# read data
films <- read.csv("dataset9.csv", header = TRUE)
# View data structures and variable types
str(films)
# View variable distribution
summary(films)
# Correlation matrix
cor(films[,c("year", "length", "budget", "votes", "rating")])
# Check for missing values
sum(is.na(films))
# Check which column has missing values
colSums(is.na(films))
# ########Fill missing values (any other way?)
mean_length <- mean(films$length, na.rm = TRUE)
films$length[is.na(films$length)] <- mean_length
# Check if missing values were filled successfully
sum(is.na(films))
```

# Step 2: Data Visualizations

```{r visualization1.1, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, out.width='90%',fig.align='center',fig.cap="\\label{fig:box1}Armed forces share population by Country.",fig.pos='H'}
# Draw a histogram (show the number of movies with different ratings, and the color distinguishes the fuzzy interval with a score of about 7)
ggplot(films, aes(x = rating, fill = factor(rating > 7))) +
  geom_histogram(alpha = 0.5, binwidth = 0.5) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9"), 
                    name = "Rating > 7", 
                    labels = c("No", "Yes")) +
  labs(title = "Distribution of IMDB ratings", x = "IMDB rating", y = "Frequency") +
  theme_classic()
```

```{r visualization1.2, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, out.width='90%',fig.align='center',fig.cap="\\label{fig:scat1}Armed forces share population by Terrorism fatalities and country,The points have been jittered.",fig.pos='h'}
# Draw a scatter plot (show the relationship between budget, ratingg and genre)
ggplot(films, aes(x = budget, y = rating)) +
  geom_point(aes(size = votes, color = genre)) +
  scale_color_brewer(palette = "Set2", name = "Genre") +
  labs(title = "Relationship between budget, rating and genre", 
       x = "Budget", y = "IMDB rating") +
  theme_classic()
```


```{r visualization2.1, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, out.width='90%',fig.align='center',fig.cap="\\label{fig:box2}Armed forces share population by century.",fig.pos='h' }
# Draw a line chart by year (the mean and median of each year are drawn in the red and blue lines)
ggplot(films, aes(x = year, y = rating, group = 1)) +
  stat_summary(fun = mean, geom = "line", color = "red", size = 1) +
  stat_summary(fun = median, geom = "line", color = "blue", size = 1) +
  scale_x_continuous(breaks = seq(1930, 2020, by = 10), 
                     labels = seq(1930, 2020, by = 10)) +
  labs(title = "Trend of IMDB rating", x = "Year", y = "IMDB rating") +
  theme_classic()
```

# Step 3: Data Processing

```{r data processiong, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# Convert year and length to continuous variables(Because time and length can be regarded as unlimited values)
films$year <- as.numeric(as.character(films$year))
films$length <- as.numeric(as.character(films$length))
# one-hot encoding
films <- cbind(films, model.matrix(~genre-1, data=films))
# Binarize the IMDB score
films$rating2 <- ifelse(films$rating > 7, 1, 0)
```

# Step 4: Model Fit

```{r model, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# Fit the glm model on the training data
fit <- glm(rating2 ~ year + length + budget + votes + genreAction + genreAnimation+genreComedy + genreDocumentary + genreDrama + genreRomance, data=train_data, family=binomial)
```

# Step 5: Variables Selection and Models Selection

Due to the P-value of previous model result, we could delect eht insignificant variables and fit a new model. And we could alse use odds ratio to choose variables, and use these new variables to fit. Then we compare with AICs belong to different models. 

```{r selection, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# After selecting variables:
# By the result of glm fit summary, we choose the variables with p-value lower than 0.05
# By the result of odds ratio, we choose the variables that have influence on rating
fit2 = glm(rating2 ~ length + budget + votes + genreAction + genreAnimation + genreDrama + genreRomance, data=films, family=binomial)
fit3 = glm(rating2 ~ length + budget + genreComedy + genreDocumentary , data=films, family=binomial)
table = matrix(c(fit$aic, fit2$aic, fit3$aic), dimnames = list(c("fit","fit2","fit3"), c("AIC")))
table
```

```{r plot2, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# Store the variables that need to be scatter plotted in a data frame
variables <- c("year", "length", "budget", "votes", "genreAction", "genreAnimation", "genreComedy", "genreDocumentary", "genreDrama", "genreRomance")
plot_data <- films[variables]

# Plot a scatterplot matrix
pairs(plot_data)
```
It can be seen from the scatter plot matrix that the scatter plots between length and year, votes and year, length and votes, budget and length, budget and votes all present a triangular shape, that is, there may be interactions, so these five Interaction terms are put into the model for comparison

```{r selection2, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#test interaction
films$year_length <- films$year * films$length
films$year_votes <- films$year * films$votes
films$length_votes <- films$length * films$votes
films$budget_length <- films$budget * films$length
films$budget_votes <- films$budget * films$votes

fit3 <- glm(rating2 ~ year + length + budget + votes + genreAction + genreAnimation+genreComedy + genreDocumentary + genreDrama + genreRomance, data=films, family=binomial)
fit4 <- glm(rating2 ~ year + length + budget + votes + genreAction + genreAnimation+genreComedy + genreDocumentary + genreDrama + genreRomance + year_length + year_votes + length_votes+budget_length+budget_votes, data=films, family=binomial)
anova(fit3, fit4, test="Chisq")
```
Model 2 adds five interaction items on the basis of Model. From the results, the residual deviation (Resid. Dev) of Model 2 is 1514.9, which is smaller than the residual deviation (Resid. Dev) of Model 1549.1, which shows that the fitting effect of Model 2 is better. At the same time, the four new interaction items in Model 2 are statistically significant, that is, their impact on the dependent variable rating2 is statistically significant.

# Step 6: odds ratio

```{r odds ratio, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# Split data into training and testing sets
set.seed(123)
train_index <- sample(1:nrow(films), size = round(nrow(films) * 0.7), replace = FALSE)
train_data <- films[train_index, ]
test_data <- films[-train_index, ]
# Predict the test data using the fitted model
test_data$predicted <- predict(fit4, newdata = test_data, type = "response")
# Calculate the ROC curve and AUC

roc_data <- roc(test_data$rating2, test_data$predicted)
roc_auc <- auc(roc_data)
roc_auc
```

```{r odds ratio1, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# Plot the ROC curve
plot(roc_data, main = "ROC Curve for Movie Rating Model",
     xlab = "False Positive Rate", ylab = "True Positive Rate", 
     print.auc = TRUE, auc.polygon = TRUE, grid=c(0.2, 0.2), col = "darkblue")
abline(a = 0, b = 1, lty = 2, col = "gray")
```
This result shows that the model works well for predicting the classification of observations in the test dataset, because the area under the ROC curve is 0.9499, which means that the model can distinguish positive and negative examples to a large extent
```{r result, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#View model results
summary(fit4)
```

```{r result2, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
# Extract Coefficients and Standard Errors
coef_df <- tidy(fit4, exponentiate = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate_if(is.numeric, list(~ round(., 2))) %>%
  mutate(lower = estimate - 1.96 * std.error,
         upper = estimate + 1.96 * std.error)

# plot the coefficients
ggplot(coef_df, aes(x = term, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +
  labs(title = "Effect of predictors on movie rating",
       x = "", y = "Odds Ratio") +
  theme_minimal()
```

According to the results, the following conclusions can be drawn:
Intercept: The log odds of the response variable being 1 when all independent variables are 0 in this model is 8.276.

Year: For each increase of 1 year, the log odds of a movie being classified as rating2=1 will decrease by 2.145e-03 units.

Length: For each increase of 1 minute in movie length, the log odds of a movie being classified as rating2=1 will decrease by 0.2969 units, with a p-value of 0.059, which is not significant enough.

Budget: For each increase of 1 unit in budget, the log odds of a movie being classified as rating2=1 will increase by 0.2304 units, with a p-value less than 0.001, which is significant.

Votes: For each increase of 1 vote, the log odds of a movie being classified as rating2=1 will decrease by 0.0043 units, with a p-value of 0.306, which is not significant enough.

genreAction, genreAnimation, genreComedy, genreDocumentary, and genreDrama: Compared to genreThriller, the log odds of movies of other genres being classified as rating2=1 are -4.126, -4.889, -0.6432, 1.447, and -5.656 units, respectively, with p-values less than 0.05, which are significant.

genreRomance: Compared to genreThriller, the log odds of romance movies being classified as rating2=1 is -4.992 units, with a p-value less than 0.001, which is significant.

year_length: For each increase of 1 year and 1 minute in movie length, the log odds of a movie being classified as rating2=1 will increase by 9.426e-05 units, with a p-value of 0.236, which is not significant enough.

year_votes: For each increase of 1 year and 1 vote, the log odds of a movie being classified as rating2=1 will increase by 2.205e-06 units, with a p-value of 0.296, which is not significant enough.

length_votes: For each increase of 1 minute in movie length, the log odds of a movie being classified as rating2=1 will increase by 1.122e-06 units, with a p-value of 0.125, which is not significant enough.

budget_length: For each increase of 1 unit in budget and 1 minute in movie length, the log odds of a movie being classified as rating2=1 will increase by 0.003978 units, with a p-value less than 0.001, which is significant.

budget_votes: For each increase of 1 unit in budget and 1 vote, the log odds of a movie being classified as rating2=1 will decrease by 1.411e-05 units, with a p-value of 0.
